### 2014-03-17 docs/vedlegg.rmd

Title
========================================================

Mer forklaring om artiklene referert til i hovedartikkelen. Spesielt de setningene som ComBat bruk er beskrevet.

## References

Combat org
http://biostatistics.oxfordjournals.org/content/8/1/118.abstract
Adjusting batch effects in microarray expression data using empirical Bayes methods.
Johnson WE, Li C, Rabinovic A.
Biostatistics. 2007 Jan;8(1):118-27. Epub 2006 Apr 21.
PMID: 16632515 [PubMed - indexed for MEDLINE] Free Article

1
Molecular profiling of single Sca-1+/CD34+,- cells--the putative murine lung stem cells.
Hittinger M, Czyz ZT, Huesemann Y, Maneck M, Botteron C, Kaeufl S, Klein CA, Polzer B.
PLoS One. 2013 Dec 31;8(12):e83917. doi: 10.1371/journal.pone.0083917. eCollection 2013.
PMID: 24391845 [PubMed - in process] 
http://www.plosone.org/article/info%3Adoi%2F10.1371%2Fjournal.pone.0083917
" Based on the hybridization date the data set is separated into two main batches of microarrays (early 2007/2008 and late 2009), which degrade further into small sub-batches. We used the ComBat algorithm [15] to adjust for the main batches. "

2
Drinking-water arsenic exposure modulates gene expression in human lymphocytes from a U.S. population.
Andrew AS, Jewell DA, Mason RA, Whitfield ML, Moore JH, Karagas MR.
Environ Health Perspect. 2008 Apr;116(4):524-31. doi: 10.1289/ehp.10861.
PMID: 18414638 [PubMed - indexed for MEDLINE] Free PMC Article
http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2290973/
"followed by empirical Bayes adjustment procedures implemented on the R platform, as described previously (Johnson et al. 2007)"

3
Epigenetic predictor of age.
Bocklandt S, Lin W, Sehl ME, Sánchez FJ, Sinsheimer JS, Horvath S, Vilain E.
PLoS One. 2011;6(6):e14821. doi: 10.1371/journal.pone.0014821. Epub 2011 Jun 22.
PMID: 21731603 [PubMed - indexed for MEDLINE] Free PMC Article
http://www.plosone.org/article/info%3Adoi%2F10.1371%2Fjournal.pone.0014821
"Batch effect were removed using the Combat algorithm [9], and one outlier sample was removed."

4.
Childhood maltreatment is associated with distinct genomic and epigenetic profiles in posttraumatic stress disorder.
Mehta D, Klengel T, Conneely KN, Smith AK, Altmann A, Pace TW, Rex-Haffner M, Loeschner A, Gonik M, Mercer KB, Bradley B, Müller-Myhsok B, Ressler KJ, Binder EB.
Proc Natl Acad Sci U S A. 2013 May 14;110(20):8302-7. doi: 10.1073/pnas.1217750110. Epub 2013 Apr 29.
PMID: 23630272 [PubMed - indexed for MEDLINE] Free PMC Article
http://www.pnas.org/content/110/20/8302.full
"To correct for confounding as a result of batch effects, the data were normalized using an empirical Bayes method for batch correction " (expression)
"Hybridization and chip batch effects were accounted for using an empirical Bayes method (43)." (methylation)

5.
Patterns of regulatory activity across diverse human cell types predict tissue identity, transcription factor binding, and long-range interactions.
Sheffield NC, Thurman RE, Song L, Safi A, Stamatoyannopoulos JA, Lenhard B, Crawford GE, Furey TS.
Genome Res. 2013 May;23(5):777-88. doi: 10.1101/gr.152140.112. Epub 2013 Mar 12.
PMID: 23482648 [PubMed - indexed for MEDLINE] Free PMC Article
http://genome.cshlp.org/content/23/5/777.full
"We improved on the previously published open chromatin measurements by accounting for batch affects that grouped the data by laboratory rather than by biological signal (see Methods). We used ComBat (Johnson et al. 2007) to remove these batch effects, after which both the DNase I and expression data clustered according to expected biological relationships (Supplemental Fig. S1)."
"Counts were quantile-normalized and scaled, and protocol batch effects were corrected using ComBat (Supplemental Fig. S1; Johnson et al. 2007)."
"Artifactual differences between protocols (Figure S1) were corrected using ComBat (Johnson et al. 2007). Because the data was too large for the ComBat model, we divided it into subsets of about 300,000 DHSs and corrected the batch affect individually on each subset. "
"To make the arrays comparable, we used ComBat to correct for this batch affect (Johnson et al. 2007). After correction with ComBat, the arrays grouped according to expected biological similarity (Figure S1). "

6.
A recurrent inactivating mutation in RHOA GTPase in angioimmunoblastic T cell lymphoma.
Yoo HY, Sung MK, Lee SH, Kim S, Lee H, Park S, Kim SC, Lee B, Rho K, Lee JE, Cho KH, Kim W, Ju H, Kim J, Kim SJ, Kim WS, Lee S, Ko YH.
Nat Genet. 2014 Mar 2. doi: 10.1038/ng.2916. [Epub ahead of print]
PMID: 24584070 [PubMed - as supplied by publisher]
http://www.nature.com/ng/journal/vaop/ncurrent/pdf/ng.2916.pdf
http://www.nature.com/ng/journal/vaop/ncurrent/full/ng.2916.html
http://www.nature.com/ng/journal/vaop/ncurrent/extref/ng.2916-S1.pdf
"We combined the RNA-seq and microarray data using the Combat algorithm36 in the Bioconductor sva package to remove batch biases due to platform differences"

A recurrent inactivating mutation in RHOA GTPase in angioimmunoblastic T cell lymphoma.
Yoo HY, Sung MK, Lee SH, Kim S, Lee H, Park S, Kim SC, Lee B, Rho K, Lee JE, Cho KH, Kim W, Ju H, Kim J, Kim SJ, Kim WS, Lee S, Ko YH.
Nat Genet. 2014 Mar 2. doi: 10.1038/ng.2916. [Epub ahead of print]
PMID: 24584070 [PubMed - as supplied by publisher]

## anbefaler ComBat

10
Batch correction of microarray data substantially improves the identification of genes differentially expressed in rheumatoid arthritis and osteoarthritis.
Kupfer P, Guthke R, Pohlers D, Huber R, Koczan D, Kinne RW.
BMC Med Genomics. 2012 Jun 8;5:23. doi: 10.1186/1755-8794-5-23.
PMID: 22682473 [PubMed - indexed for MEDLINE] Free PMC Article
Related citations
http://www.biomedcentral.com/1755-8794/5/23

11
Relative impact of key sources of systematic noise in Affymetrix and Illumina gene-expression microarray experiments.
Kitchen RR, Sabine VS, Simen AA, Dixon JM, Bartlett JM, Sims AH.
BMC Genomics. 2011 Dec 1;12:589. doi: 10.1186/1471-2164-12-589.
PMID: 22133085 [PubMed - indexed for MEDLINE] Free PMC Article
http://www.biomedcentral.com/1471-2164/12/589/

Tackling the widespread and critical impact of batch effects in high-throughput data.
Leek JT, Scharpf RB, Bravo HC, Simcha D, Langmead B, Johnson WE, Geman D, Baggerly K, Irizarry RA.
Nat Rev Genet. 2010 Oct;11(10):733-9. doi: 10.1038/nrg2825. Epub 2010 Sep 14.
PMID: 20838408 [PubMed - indexed for MEDLINE] Free PMC Article
http://www.nature.com/nrg/journal/v11/n10/abs/nrg2825.html

A comparison of batch effect removal methods for enhancement of prediction performance using MAQC-II microarray gene expression data.
Luo J, Schumacher M, Scherer A, Sanoudou D, Megherbi D, Davison T, Shi T, Tong W, Shi L, Hong H, Zhao C, Elloumi F, Shi W, Thomas R, Lin S, Tillinghast G, Liu G, Zhou Y, Herman D, Li Y, Deng Y, Fang H, Bushel P, Woods M, Zhang J.
Pharmacogenomics J. 2010 Aug;10(4):278-91. doi: 10.1038/tpj.2010.57.
PMID: 20676067 [PubMed - indexed for MEDLINE] Free PMC Article
http://www.nature.com/tpj/journal/v10/n4/full/tpj201057a.html

Removing batch effects in analysis of expression microarray data: an evaluation of six batch adjustment methods.
Chen C, Grennan K, Badner J, Zhang D, Gershon E, Jin L, Liu C.
PLoS One. 2011 Feb 28;6(2):e17238. doi: 10.1371/journal.pone.0017238.
PMID: 21386892 [PubMed - indexed for MEDLINE] Free PMC Article
http://www.plosone.org/article/info%3Adoi%2F10.1371%2Fjournal.pone.0017238

T. Hastie, R. Tibshirani, Balasubramanian Narasimhan and Gil Chu (2013). pamr: Pam: prediction analysis for microarrays. R package version
  1.54.1. http://CRAN.R-project.org/package=pamr

## skeptisk til ComBat

20 - HAr kjoert permutering av covariatelabler og finner at combat finner masse gener uansett. Fokus paa artikkel er ikke combat spesifikkt.
Expanding the understanding of biases in development of clinical-grade molecular signatures: a case study in acute respiratory viral infections.
Lytkin NI, McVoy L, Weitkamp JH, Aliferis CF, Statnikov A.
PLoS One. 2011;6(6):e20662. doi: 10.1371/journal.pone.0020662. Epub 2011 Jun 1.
PMID: 21673802 [PubMed - indexed for MEDLINE] Free PMC Article
http://www.plosone.org/article/info%3Adoi%2F10.1371%2Fjournal.pone.0020662

## ComBat i pipelines

TCGA Batch Effects Tool
http://bioinformatics.mdanderson.org/main/TCGABatchEffects:Overview

GenePattern 2.0.
Reich M, Liefeld T, Gould J, Lerner J, Tamayo P, Mesirov JP.
Nat Genet. 2006 May;38(5):500-1. No abstract available.
PMID: 16642009 [PubMed - indexed for MEDLINE]
http://www.nature.com/ng/journal/v38/n5/full/ng0506-500.html

Uverifisert.
insilicodb? bioconductopakke eller hva
http://bib.oxfordjournals.org/content/early/2012/07/31/bib.bbs037.short

The sva package for removing batch effects and other unwanted variation in high-throughput experiments.
Leek JT, Johnson WE, Parker HS, Jaffe AE, Storey JD.
Bioinformatics. 2012 Mar 15;28(6):882-3. doi: 10.1093/bioinformatics/bts034. Epub 2012 Jan 17.
PMID: 22257669 [PubMed - indexed for MEDLINE] Free PMC Article
http://bioinformatics.oxfordjournals.org/cgi/pmidlookup?view=long&pmid=22257669

virtualArray: a R/bioconductor package to merge raw data from different microarray platforms.
Heider A, Alt R.
BMC Bioinformatics. 2013 Mar 2;14:75. doi: 10.1186/1471-2105-14-75.
PMID: 23452776 [PubMed - indexed for MEDLINE] Free PMC Article
http://bioconductor.org/packages/release/bioc/html/virtualArray.html

inSilicoDb: an R/Bioconductor package for accessing human Affymetrix expert-curated datasets from GEO.
Taminau J, Steenhoff D, Coletta A, Meganck S, Lazar C, de Schaetzen V, Duque R, Molter C, Bersini H, Nowé A, Weiss Solís DY.
Bioinformatics. 2011 Nov 15;27(22):3204-5. doi: 10.1093/bioinformatics/btr529. Epub 2011 Sep 21.
PMID: 21937664 [PubMed - indexed for MEDLINE] Free Article

## eksempler med reporduserte p-value

Comparing the biological impact of glatiramer acetate with the biological impact of a generic.
Towfic F, Funt JM, Fowler KD, Bakshi S, Blaugrund E, Artyomov MN, Hayden MR, Ladkani D, Schwartz R, Zeskind B.
PLoS One. 2014 Jan 8;9(1):e83757. doi: 10.1371/journal.pone.0083757. eCollection 2014.
PMID: 24421904 [PubMed - in process] Free PMC Article
http://www.plosone.org/article/info%3Adoi%2F10.1371%2Fjournal.pone.0083757

### user group recomandations
https://groups.google.com/d/msg/combat-user-forum/eSVSKwGtuyE/ZIWV2juYmmAJ
https://groups.google.com/d/msg/combat-user-forum/Vkb9p7wekd4/h5Etie7FTVQJ
https://groups.google.com/d/msg/combat-user-forum/fpBTcgDjiR8/uo4QIZL4sZgJ
https://groups.google.com/d/msg/combat-user-forum/26FZlgU2LFQ/W6U_Lhh_64EJ