### 2014-03-17 reanalysis/test/batch_adjust_no_covariates_can_induce_group_diff.rmd

========================================================

Show that adjusting for batch without trying to retain group differeces also can induce group effects.

```{r}
library(sva)
source("../../commonscripts/helper_functions.r")
```

Creating a unbalances experiment, 3 conditions, 2 batches with some genes true different in condition 3 compared to 1 and 2. But 1 and 2 have no difference. No batch effect added but adjusted for anyways.
```{r}
n=10
ngenes=5000
ndiffgenes = round(ngenes/10)
sampleannotation = createsampleannotation(  list(c(n,n,0), c(n,0,n))) 
#sampleannotation = createsampleannotation(  list(c(n,1), c(1,1), c(1,n))) 
randomdata = createrandomdata(ngenes, sampleannotation, mean=0, sd=1)
conditiondata = addconditioneffect(df=randomdata, labels=sampleannotation$treatment, ndiffgenes=ndiffgenes, thismean=10, betweensd=1, insidesd=1, affectedconditions="treatment3")
print(table(sampleannotation[, c("batch", "treatment")]))
```

Then run ComBat without covariates.
```{r}
batcnnormdata = ComBat(dat=conditiondata, 
                       batch=sampleannotation$batch, 
                       mod=NULL, par.prior=TRUE, prior.plots=FALSE)

```

And plot p-values
```{r pvaluespermuted, dev='svg', fig.width=10, fig.height=10, tidy=FALSE}
ret=getdifftab_limma(batcnnormdata, sampleannotation$treatment, "treatment1-treatment2")
#table(ret$padjusted<0.05)
hist(ret$p, breaks=100, border="red")

ret=getdifftab_limma(conditiondata, sampleannotation$treatment, "treatment1-treatment2", block=sampleannotation$batch)
#table(ret$padjusted<0.05)
hist(ret$p, breaks=100, add=T, border="black")

a = sampleannotation$batch==1
ret=getdifftab_limma(conditiondata[,a], sampleannotation$treatment[a], "treatment1-treatment2")
#table(ret$padjusted<0.05)
hist(ret$p, breaks=100, add=T, border="blue")
legend("topright", legend=c("ComBat adjusted", "Limma adjusted", "Only batch 1, no adjustment"), text.col=c("red", "black", "blue"))

```
Treatment1 and Treatment2 differs. Some of the effect in Treatment3 has been transfered overt to Treatment1 since they both are in batch2.
This problem is not limited to ComBat.

```{r}
library(pamr)

pamrnnormdata = pamr.batchadjust(list(x=conditiondata,
                                      y=factor(sampleannotation$treatment),
                                      batchlabels=factor(sampleannotation$batch)))
ret=getdifftab_limma(pamrnnormdata[["x"]], sampleannotation$treatment, "treatment1-treatment2")
#table(ret$padjusted<0.05)
hist(ret$p, breaks=100, border="red")

```